---
title: "Frequency effects on duration in three langauges"
output: html_document
---
```{r setup, include=FALSE, echo=FALSE}
library(here)
library(readr)
library(tidyverse)
library(brms)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(bayestestR)
```

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
# load needed data 
span_data <- read.csv(here("corpus_data", "tidy", "span_tidy.csv"))
eng_data <- read.csv(here("corpus_data", "tidy", "eng_tidy.csv"))
germ_data <- read.csv(here("data", "tidy", "german_tidy.csv")) 
span_model  <- read_rds(here("corpus_data", "models", "span_model_bayes.RDS"))
eng_model <- read_rds(here("corpus_data", "models", "eng_model_bayes.RDS"))
german_model <- read_rds(here("data", "models", "german_model.rds"))

fixef_df_span <- fixef(span_model) %>% 
  as.data.frame()
fixef_df_eng <- fixef(eng_model) %>% 
  as.data.frame()
fixef_df_germ <- fixef(german_model) %>% 
  as.data.frame()

span_model_df <- span_model %>% 
  as.data.frame()
hdi_span <- hdi(span_model_df$b_log_freq_z) 

# calculate range of difference from most to least frequent tokens 
low = (fixef(span_model)[2,1]*min(span_data$log_freq, na.rm=TRUE) + fixef(span_model)[1,1])*1000 %>% 
  as.data.frame()
high = (fixef(span_model)[2,1]*max(span_data$log_freq, na.rm=TRUE) + fixef(span_model)[1,1])*1000 %>% 
  as.data.frame()
low_e = (fixef(eng_model)[2,1]*min(eng_data$log_freq, na.rm=TRUE) + fixef(eng_model)[1,1])*1000 %>% 
  as.data.frame()
high_e = (fixef(eng_model)[2,1]*max(eng_data$log_freq, na.rm=TRUE) + fixef(eng_model)[1,1])*1000 %>% 
  as.data.frame()

low_obs_e <- min(eng_data$log_freq, na.rm=TRUE)
slope_e <- fixef(eng_model)[2,1]
int_e <- fixef(eng_model)[1,1]

low_obs_s <- min(span_model$log_freq, na.rm=TRUE)
slope_s <- fixef(span_model)[2,1]
int_s <- fixef(span_model)[1,1]
# when x = 10.27, y = .559
# what does y equal when x = 0?
# multiple low_obs by slope minus intercept 
y_int_e <- (low_obs_e*slope_e)*-1 + int_e
y_int_s <- (slope_s*slope_s)*-1 + int_s

fixef_df_eng <- fixef(eng_model) %>% 
  as.data.frame()
eng_model_df <- eng_model %>% 
  as.data.frame()
hdi_eng <- hdi(eng_model_df$b_log_freq) 

corpus <- read.csv(here("lexicalfrequencydata", "subtlex_tidy.csv"))
freq <- read.csv(here("eng_corpus_data", "tidy", "unigram_freq.csv"))

# for plots 
span_post <- as_tibble(span_model) %>% select(int = b_Intercept, b = b_log_freq_z)
eng_post <- as_tibble(eng_model) %>% select(int = b_Intercept, b = b_log_freq_z) 

germ_post <- as_tibble(german_model) %>% select(int = b_Intercept, b = b_log_freq_z)

```


##### Figure 1: Spanish data and most plausible line of best fit
```{r}
# spanish plot 
span_data %>% 
  ggplot(aes(x = duration_z, y = log_freq_z, color = length_z)) + geom_point() + geom_abline(intercept = 
                                                                            fixef(span_model)[1, 1],
                                                                           slope =  fixef(span_model)[2, 1], color = "darkred", size = 1.5)
                       
```


##### Figure 2: English data and most plausible line of best fit
```{r}

eng_data %>% 
  ggplot(aes(x = duration_z, y = log_freq_z, color = length_z)) + geom_point() + geom_abline(intercept = 
                                                                            fixef(span_model)[1, 1],
                                                                           slope =  fixef(span_model)[2, 1], color = "darkred", size = 1.5)
```


##### Figure 3: German data and most plausible line of best fit
```{r}

germ_data %>% 
  ggplot(aes(x = duration_z, y = log_freq_z, color = length_z)) + geom_point() + geom_abline(intercept = 
                                                                            fixef(german_model)[1, 1],
                                                                           slope =  fixef(german_model)[2, 1], color = "darkred", size = 1.5)

```

##### Table 1: Spanish model 
```{r}
fixef(span_model) %>% 
  knitr::kable(digits = 3)
```

##### Table 2: English model 
```{r}
fixef(eng_model) %>% 
  knitr::kable(digits = 3)
```

##### Table 3: German model 
```{r}
fixef(german_model) %>% 
  knitr::kable(digits = 3)
```


##### Figure 4: English Posterior distributions with medians and 80% intervals
```{r}
plot_title_eng <- ggtitle("English Posterior distributions",
                      "with medians and 80% intervals")
bayesplot::mcmc_areas(eng_post,
           pars = c("b"),
           prob = 0.8) + geom_vline(xintercept = 0) + xlim(-1, 1)
```

##### Figure 5: Spanish Posterior distributions with medians and 80% intervals
```{r}
plot_title_span <- ggtitle("Spanish Posterior distributions",
                      "with medians and 80% intervals")
bayesplot::mcmc_areas(span_post,
           pars = c("b"),
           prob = 0.8) + geom_vline(xintercept = 0) + xlim(-1, 1)
```

##### Figure 6: German Posterior distributions with medians and 80% intervals
```{r}
plot_title_german <- ggtitle("German Posterior distributions",
                      "with medians and 80% intervals")
bayesplot::mcmc_areas(germ_post,
           pars = c("b"),
           prob = 0.8) + geom_vline(xintercept = 0) + xlim(-1, 1)
```