---
title: "The effect of lexical frequency on word duration: analyzing corpus data in Spanish and English"
author: ""
date: ""
output: word_document
#  pdf_document: 
#    fig_caption: yes
#    keep_tex: yes
header-includes:
   - \usepackage{tipa}
   - \usepackage{xcolor}
fontsize: 12pt
#csl: ../../bib/apa.csl

bibliography      : 
  - "./bib/references.bib"
---

```{r setup, include = FALSE}
library("papaja")
library("knitr")
source(here::here("scripts", "00_libraries.R"))
```

The present study investigates the effect of lexical frequency on word duration in Spanish and English monolingual speech. 
Previous studies have found that word duration shortens in English in more frequent words [e.g., @gahl_2008; @gahl_2009; @lohmann_2018]. 
The shortening of frequent forms may correspond to articulatory routinization [@bybee_2001; @newmeyer_2006], but evidence showing that in homophone pairs (e.g., *thyme* – *time*), the item with higher frequency is shorter than the infrequent one reveals that frequency effects on duration may not be due to repetition of a phonological form but to lemma frequency effects instead [@gahl_2008; @gahl_2009]. 

The present study aims to replicate the frequency effect found in duration in English [@gahl_2008; @gahl_2009; @lohmann_2018] and explore the effect of lexical frequency on word duration in Spanish. 
This study analyzes corpus data, with the English data coming from the Free ST American English Corpus and Spanish data from the Crowdsourced high-quality Argentinian Spanish speech data set [@guevara-rukoz-etal-2020-crowdsourcing].
The English data were 2806 recordings of cellphone conversations, and the Spanish data consisted of 1928 recordings of random Spanish sentences recorded by volunteers in Buenos Aires, Argentina.
Lexical frequency was measured using movie subtitle frequencies [see @new_use_2007].
The data was analyzed using two Bayesian linear regressions with duration as the outcome variable and speech rate and orthographic length as fixed predictors. 

The results replicated the frequency effects previously found in English, and expanded them to Spanish. 
English frequent words were found to be shorter than infrequent ones when orthographic length and speech rate were controlled for (Figure 1).
We found similar results in Spanish, where the results exhibited a negative linear relationship between lexical frequency and word duration (Figure 1). 
The findings have implications for neo-generative [@levelt_1989], exemplar [@johnson_2006], and hybrid [@pierrehumbert_2016] models of sound representation. 

\clearpage

```{r, 'fig-cap', echo=F}
fig_caption <- "Whole word duration in English (Left Panel) and Spanish (Right Panel) as a function of lexical frequency and orthographic length (length_z) with the most plausible line of best fit."
```

```{r, 'plot-joined', echo=F, fig.cap=fig_caption, out.width="100%"}
knitr::include_graphics(
  here("docs", "abstracts", "NewSounds22", "figs", "joined_plots.png")
)
```

\begin{center}
References
\end{center}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\phantom{.} 
\textcolor{white}{\\} 
\vspace{-0.5in}

<div id = "refs"></div>
\endgroup